PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-04 paper/app-pineappl.tex
Add more smaller changes

\lstset{
  basicstyle=\ttfamily\small,
  showstringspaces=false,
  commentstyle=\color{red},
  keywordstyle=\color{blue}
}

@@

\paragraph{Channels and Luminosity function.}
Sometimes it is usuful to know which partons contribute the most and by how much.
This is what the subcommand \texttt{channels} shows:
\begin{verbatim}
 $ pineappl channels ATLASZHIGHMASS49FB.pineappl \
 > NNPDF31_nlo_as_0118_luxqed --limit 5
bin xmin xmax lumi  size
---+----+----+----+------+---+------+---+------+---+------+---+-----
  0  116  130  #15 27.42%  #0 27.42%  #5 24.52% #20 24.51% #30 1.30%
  1  130  150  #15 28.83%  #0 28.83%  #5 21.78% #20 21.78% #30 1.86%
  2  150  170  #15 29.87%  #0 29.82%  #5 19.63% #20 19.62% #30 2.32%
  3  170  190  #15 30.54%  #0 30.53% #20 18.17%  #5 18.13% #30 2.58%
  4  190  210   #0 31.06% #15 31.04%  #5 17.07% #20 17.04% #30 2.75%
  5  210  230   #0 31.46% #15 31.38% #20 16.27%  #5 16.24% #30 2.86%
  6  230  250   #0 31.73% #15 31.64% #20 15.66%  #5 15.62% #30 2.91%
  7  250  300   #0 32.07% #15 32.02%  #5 14.91% #20 14.86% #30 2.98%
  8  300  400   #0 32.62% #15 32.56%  #5 13.83% #20 13.81% #30 2.98%
  9  400  500  #15 33.14%  #0 33.12% #20 12.89%  #5 12.87% #30 2.93%
 10  500  700   #0 33.71% #15 33.31% #20 12.28%  #5 12.26% #30 2.93%
 11  700 1000  #15 34.08%  #0 33.71% #20 11.59%  #5 11.51% #30 3.06%
 12 1000 1500   #0 33.95% #15 33.88% #20 11.12%  #5 10.88% #30 3.56%
\end{verbatim}
The first three columns are known from \texttt{convolute}.
The next columns (the switch \texttt{-{}-limit 5} limits the output to five columns) show first the channel index and then relative size of the corresponding contribution.
Since the contribution of a partonic channel can be negative, the columns are sorted ignoring the sign of the contribution.
The first line shows that for bin 0, i.e.\ for the range $\SI{116}{\giga\electronvolt} < M_{\ell \bar{\ell}} < \SI{130}{\giga\electronvolt}$, the cross section is dominated by partonic channel \texttt{\#15} (\SI{27.42}{\percent}), following by partonic channel \texttt{\#0}, then \texttt{\#5}, etc.
The meaning of the channel numbers is given by using the subcommand \texttt{luminosity} (only an excerpt is shown):
\begin{verbatim}
id    entry
--+------------+------------+------------
0  1 × ( 2, -2) 1 × ( 4, -4)
5  1 × ( 1, -1) 1 × ( 3, -3)
15 1 × (-4,  4) 1 × (-2,  2)
20 1 × (-3,  3) 1 × (-1,  1)
30 1 × (22, 22)
\end{verbatim}
This shows that channel \texttt{\#0} represents the up-type quark--anti-quark contribution (shown with PDG id 2 and 4 for up and charm quarks, which have the same matrix elements), channel \texttt{\#15} is the same channel with initial states transposed, channels \texttt{\#5} and \texttt{\#20} are the down-type quark--anti-quark channels, and channel \texttt{\#30} is the photon--photon channel.
The factors \texttt{1} are not important here, but in general they can contain CKM matrix elements that would otherwise not allow to share the matrix elements between different quark flavours.

@@

\lstset{
  basicstyle=\ttfamily\small,
  showstringspaces=false,
  commentstyle=\ttfamily\small,
  keywordstyle=\ttfamily\small,
  identifierstyle=\color{black},
  stringstyle=\color{black},
}

>>>>

\lstset{
  basicstyle=\ttfamily\small,
  showstringspaces=false,
  commentstyle=\color{red},
  keywordstyle=\color{blue},
}

@@

\paragraph{Channels and Luminosity function.}
Sometimes it is useful to know which partons contribute the most and by how much.
This is what the subcommand \texttt{channels} shows:
\begin{verbatim}
 $ pineappl channels ATLASZHIGHMASS49FB.pineappl \
 > NNPDF31_nlo_as_0118_luxqed --limit 5
bin xmin xmax lumi  size
---+----+----+----+------+---+------+---+------+---+------+---+-----
  0  116  130  #15 27.42%  #0 27.42%  #5 24.52% #20 24.51% #30 1.30%
  1  130  150  #15 28.83%  #0 28.83%  #5 21.78% #20 21.78% #30 1.86%
  2  150  170  #15 29.87%  #0 29.82%  #5 19.63% #20 19.62% #30 2.32%
  3  170  190  #15 30.54%  #0 30.53% #20 18.17%  #5 18.13% #30 2.58%
  4  190  210   #0 31.06% #15 31.04%  #5 17.07% #20 17.04% #30 2.75%
  5  210  230   #0 31.46% #15 31.38% #20 16.27%  #5 16.24% #30 2.86%
  6  230  250   #0 31.73% #15 31.64% #20 15.66%  #5 15.62% #30 2.91%
  7  250  300   #0 32.07% #15 32.02%  #5 14.91% #20 14.86% #30 2.98%
  8  300  400   #0 32.62% #15 32.56%  #5 13.83% #20 13.81% #30 2.98%
  9  400  500  #15 33.14%  #0 33.12% #20 12.89%  #5 12.87% #30 2.93%
 10  500  700   #0 33.71% #15 33.31% #20 12.28%  #5 12.26% #30 2.93%
 11  700 1000  #15 34.08%  #0 33.71% #20 11.59%  #5 11.51% #30 3.06%
 12 1000 1500   #0 33.95% #15 33.88% #20 11.12%  #5 10.88% #30 3.56%
\end{verbatim}
The first three columns are known from \texttt{convolute}.
The next columns (the switch \texttt{-{}-limit 5} limits the output to five columns) show first the channel index and then relative size of the corresponding contribution.
Since the contribution of a partonic channel can be negative, the columns are sorted ignoring the sign of the contribution.
The first line shows that for bin 0, i.e.\ for the range $\SI{116}{\giga\electronvolt} < M_{\ell \bar{\ell}} < \SI{130}{\giga\electronvolt}$, the cross section is dominated by partonic channel \texttt{\#15} (\SI{27.42}{\percent}), following by partonic channel \texttt{\#0}, then \texttt{\#5}, etc.
The meaning of the channel numbers is given by using the subcommand \texttt{luminosity} (only an excerpt is shown):
\begin{verbatim}
id    entry
--+------------+------------+------------
0  1 × ( 2, -2) 1 × ( 4, -4)
5  1 × ( 1, -1) 1 × ( 3, -3)
15 1 × (-4,  4) 1 × (-2,  2)
20 1 × (-3,  3) 1 × (-1,  1)
30 1 × (22, 22)
\end{verbatim}
This shows that channel \texttt{\#0} represents the up-type quark--anti-quark contribution (shown with PDG id 2 and 4 for up and charm quarks, which have the same matrix elements), channel \texttt{\#15} is the same channel with initial states transposed, channels \texttt{\#5} and \texttt{\#20} are the down-type quark--anti-quark channels, and channel \texttt{\#30} is the photon--photon channel.
The factors \texttt{1} are not important here, but in general they can contain CKM matrix elements that would otherwise not allow to share the matrix elements between different quark flavours.

@@

### 2020-08-04 paper/sec-doublecounting.tex
Add more smaller changes

\section{On the consistency of experimental data}
\label{sec:doublecounting}

@@

A second example is related to removing EW effects from data. These can be either the full EW corrections
or just a part of them. In either case, a comparison between these data and a NLO-EW accurate simulation aimed at the extraction of some parameter would be meaningless, as some effects included in the latter
have been removed from the former. The typical example relevant for LHC is the deconvolution of effects due to multiple-photon radiation
from light particles in the final state. This applies mostly
to processes such as neutral- or charge-current Drell--Yan, specially when electrons are considered. The problem lies in the fact that
 electrons, and to a lesser extent muons, tend to radiate photons, which are not accounted
for in QCD-only matrix-elements. Thus, leptons that are measured in the detector are less energetic, and this fact is compensated for
by inverting a photon shower. The resulting dataset is e.g.\ referred to as \emph{pre-FSR} with observables defined in terms of \emph{Born-level electrons} (see e.g.\ ref.~\cite{Aad:2015auj} for its definition).
These datasets are needed for and correctly used in QCD-only PDF determinations, since the EW corrections to some DY observables can be significant, and excluding them would therefore degrade the quality of the fit.
In fits including fixed-order EW corrections the problem with this definition is that the first photon emission is included exactly at the matrix-element level. The inclusion of
subsequent emission would require the matching with the QED shower, which is not yet available for general processes.

It is interesting
to note that one can tune the QED parton shower to specific decays to mimic NLO EW effects, so that a prediction only accurate at NLO
QCD displays a remarkable agreement with another at NLO QCD+EW when the photonic shower is included (see for example the behaviour of predictions showered with
\textsc{Photos}~\cite{Barberio:1990ms,Barberio:1993qi,Golonka:2005pn} in ref.~\cite{CarloniCalame:2016ouw}). However, this kind of agreement
always comes \emph{a posteriori}, and cannot be ensured in general.
Another problem with this approach is that it introduces a dependence on the program, or even on a specific version of the program if bugs have been found.
This is especially problematic if these datasets are the only ones published, since undoing the exact subtraction can be very difficult or practically impossible as time passes (TODO insert comment about LEP ISR and DIS EW effects removal).

>>>>

\section{Subtraction of EW effects from data}
\label{sec:doublecounting}

@@

A second example is related to removing EW effects from data. These can be either the full EW corrections
or just a part of them. In either case, a comparison between these data and a NLO-EW accurate simulation aimed at the extraction of some parameter would be meaningless, as some effects included in the latter
have been removed from the former. The typical example relevant for LHC is the deconvolution of effects due to multiple-photon radiation
from light particles in the final state. This applies mostly
to processes such as neutral- or charge-current Drell--Yan, especially when electrons are considered. The problem lies in the fact that
 electrons, and to a lesser extent muons, tend to radiate photons, which are not accounted
for in QCD-only matrix-elements. Thus, leptons that are measured in the detector are less energetic, and this fact is compensated for
by inverting a photon shower. The resulting dataset is e.g.\ referred to as \emph{pre-FSR} with observables defined in terms of \emph{Born-level electrons} (see e.g.\ ref.~\cite{Aad:2015auj} for its definition).
These datasets are needed for and correctly used in QCD-only PDF determinations, since the EW corrections to some DY observables can be significant, and excluding them would therefore degrade the quality of the fit.
However, in fits including fixed-order EW corrections the problem with this definition is that the first photon emission is included exactly at the matrix-element level. The inclusion of
subsequent emission would require the matching with the QED shower, which is not yet available for general processes.

It is interesting
to note that one can tune the QED parton shower to specific decays to mimic NLO EW effects, so that a prediction only accurate at NLO
QCD displays a remarkable agreement with another at NLO QCD+EW when the photonic shower is included (see for example the behaviour of predictions showered with
\textsc{Photos}~\cite{Barberio:1990ms,Barberio:1993qi,Golonka:2005pn} in ref.~\cite{CarloniCalame:2016ouw}). However, this kind of agreement
always comes \emph{a posteriori}, and cannot be ensured in general.
Another problem with this approach is that it introduces a dependence on the program that is used for the shower inversion, or even on a specific version of the program, for examples if bugs have been found.
This is especially problematic if these datasets are the only ones published, since undoing the exact subtraction can be very difficult or practically impossible as time passes (TODO insert comment about LEP ISR and DIS EW effects removal).
